---
title: "Breast Cancer Summary"
author: "Aneesh Chandramouli, Jiangnan Lyu, Zainab Alkhater"
output: html_document
header-includes:
   - \usepackage{setspace}
   - \doublespacing
---

<br/> 

```{r message=F,echo=FALSE,results='hide',warning=FALSE}
# In this step, we renamed and recoded the chemotherapy and surgery type variables 
dat <- readRDS(
  "C:/Users/Zainab/Desktop/UM MS BIOSTAT/summar/machien/analysis_city20200601(1).rds"
  )
tidydat <- dat %>%
  rename(Chemotherapy = colnames(dat)[2]) %>%
  rename(`Surgery Type` = colnames(dat)[3]) %>%
  mutate(Chemotherapy = as_factor(Chemotherapy)) %>%
  mutate(Chemotherapy = fct_recode(Chemotherapy, 
                                   "None" = "0",
                                   "Neoadjuvant (before surgery)" = "1",
                                   "Adjuvant (after surgery)" = "2",
                                   "Only chemotherapy/ No surgery" = "3",
                                   "Recommended (unknown if received)" = "4")) %>% 
  mutate(`Surgery Type` = as_factor(`Surgery Type`)) %>%
  mutate(`Surgery Type` = fct_recode(`Surgery Type`, 
                                     "None" = "0",
                                     "Lumpectomy" = "1",
                                     "Mastectomy/MRM" = "2",
                                     "Unknown" = "3"))
```

## The variables that are missing

* Birth place
* Relationship
* Individual SES
* Neighborhood SES
* Area Deprivation Index
* Residence in Ethnic Enclave
* Education
* Tobacco Use
* Alcohol Use
* Obesity
* No Moderate Physical Activity
* Age at Menarche
* Age at First Full-Term Live Birth
* Number of Live Births
* Breastfeeding
* Personal History of Benign Breast Disease
* Personal History of Ovarian Cancer
* Family History of Breast Cancer
* Current or Prior History of Hormone Replacement Therapy (HRT)
* Comorbidities
* Receptor Status
* Treatment


<br/> 

```{r echo=FALSE}
# In this step, we renamed and recoded the final clinical stage and tumor grade variables

breastCancer <- tidydat %>%
  select(everything(), - Age) %>%
  map_df(as_factor) %>%
  mutate(
    `Final Clinical Stage` = finalClinicalStage,
    `Tumor Grade` = fct_recode(
      as_factor(tumorGrade),
      "Anaplastic" = "anaplastic",
      "Poorly differentiated / Unknown" = "poor/unknown",
      "Well/moderately differentiated" = "well/mod"),
    `Race & Ethnicity` = case_when(
    Hispanic == 0 & Race1Desc == "Black" ~ 1,
    Hispanic == 0 & Race1Desc == "White" ~ 2,
    Hispanic == 1 ~ 3),
    `Race & Ethnicity` = as_factor(`Race & Ethnicity`),
    `Race & Ethnicity` = fct_recode(`Race & Ethnicity`,
                                    "Non-Hispanic Black" = "1",
                                    "Non-Hispanic White" = "2",
                                    "Hispanic" = "3")
  )

# In this step, we created a frequency table in order to observe the frequency counts
#   and percentages of patients' final clinical stage and tumor grade by race/ethnicity
library(table1)
table1(~ `Final Clinical Stage` + `Tumor Grade` | `Race & Ethnicity`, breastCancer)
```


```{r echo=FALSE,fig.align='c'}
# Here we made both a histogram and violin plot of patients' age at diagnosis by 
#   race/ethnicity

histo_agedx_race <- 
  breastCancer %>% 
  ggplot(aes(x = tidydat$Age, fill = `Race & Ethnicity`)) + 
  geom_histogram(bins = 11) +
  labs(
    fill = "Race/Ethnicity", 
    title = "Histogram of Patients' Age at Diagnosis by Race/Ethnicity"
    ) + 
  xlab("Age at Diagnosis") + ylab("Count")
vio_agedx_race <- 
  breastCancer %>% 
  ggplot(aes(x = `Race & Ethnicity`, y = tidydat$Age)) + 
  geom_violin(aes(fill = `Race & Ethnicity`)) +
  labs(
    fill = "Race/Ethnicity", 
    title = "Violin Plot of Patients' Age at Diagnosis by Race/Ethnicity"
    ) + 
  xlab("Race/Ethnicity") + ylab("Age at Diagnosis")
```

<br/> 

```{r echo=FALSE}
histo_agedx_race
```

<br/> 

```{r echo=FALSE}
vio_agedx_race
```







