---
title: "Calculate the sum of a vector"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculate the sum of a vector}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(someWHAT)
```

# About

Function some( ) takes a vector and returns the total (what could be called the sum of the values).
Function some() checks to make sure that the user passes data that can be used for a sum.

```{r message=FALSE}
x<-round(rnorm(3,0,1),1)
x
some(x)
```

If a user does not pass some( ) valid data it will return error messages:

If the input vector contains non-numeric elements, the function execution stops and an error message "Argument `x` should be numeric" is returned.

```{r message=FALSE,error=TRUE}
x<-c("hello",1,2)
x
some(x)

```

If there is any missing elements in the vector, the function execution stops and an error message "Missing values are not allowed" is returned. 

```{r message=FALSE,error=TRUE}
x<-c(5,NA,1)
x
some(x)

```

  
